---
title: "Foraging Model"
author: "Ben Weinstein"
date: "`r Sys.time()`"
output: 
  html_document:
    toc: true
    number_sections: true
    theme: united
---

# Foraging Model

The proportion of time an animal is in a feeding behavioral state.

(@eq1)

*Process Model*

$$Y_{i,t+1} \sim Multivariate Normal(d_{i,t},σ)$$

$$d_{i,t}= Y_{i,t} + γ_{s_{i,g,t}}*T_{i,g,t}*( Y_{i,g,t}- Y_{i,g,t-1} )$$

$$
\begin{matrix}
  \alpha_{i,1,1} & \beta_{i,1,1} & 1-(\alpha_{i,1,1} + \beta_{i,1,1}) \\
  \alpha_{i,2,1} & \beta_{i,2,1} & 1-(\alpha_{i,2,1} + \beta_{i,2,1}) \\
    \alpha_{i,3,1} & \beta_{i,3,1} & 1-(\alpha_{i,3,1} + \beta_{i,3,1}) \\
\end{matrix}
$$
$$logit(\phi_{Behavior}) = \alpha_{Behavior_{t-1}} $$
The behavior at time t of individual i on track g is a discrete draw.
$$S_{i,g,t} \sim Cat(\phi_{traveling},\phi_{foraging},\phi_{resting})$$

Dive information is a mixture model based on behavior (S)

$\text{Average dive depth}(\psi)$
$$ DiveDepth \sim Normal(dive_{\mu_S},dive_{\tau_S})$$


```{r,echo=F,warning=F,message=F}
library(tidyr)
library(ggplot2)
library(maptools)
library(shiny)
library(raster)
library(data.table)
library(ggmap)
library(leaflet)
library(dplyr)
library(stringr)
library(chron)
library(R2jags)
library(boot)
library(knitr)
library(MCMCpack)
library(truncnorm)

newModel=T

opts_chunk$set(echo=F,warning=F,message=F,fig.width=11)

mytheme<-theme(axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.text.y=element_blank(),axis.ticks.y=element_blank(),axis.title.x=element_blank(),axis.title.y=element_blank(),panel.grid=element_blank())
```

```{r}
if(!newModel){
  load("/Users/Ben/Dropbox/Whales/Dive/WhalePhys.RData")
  newModel=F
}

#get gps data
f<-list.files("Data/Humpback",pattern="Locations",full.names=T,recursive = T)
gdat<-lapply(f,function(x) read.csv(x,stringsAsFactors=F))
gdat<-lapply(gdat,function(x){
  x$Quality<-as.character(x$Quality)
  return(x)
}) 

gdat<-bind_rows(gdat)

#timestamp
gdat$timestamp<-as.POSIXct(gdat$Date,format="%H:%M:%S %d-%b-%Y",tz="GMT")

#
gdat<-gdat[!is.na(gdat$Latitude),]

#crop by extent
d<-SpatialPointsDataFrame(cbind(gdat$Longitude,gdat$Latitude),data=data.frame(gdat),proj4string=CRS("+proj=longlat +datum=WGS84"))

cropoly<-readShapePoly("Data/CutPolygon.shp",proj4string=CRS("+proj=longlat +datum=WGS84"))

b<-d[!is.na(d %over% cropoly)[,2],]

gdat<-b@data

#get dive data files
f<-list.files("Data/Humpback/",pattern="Behavior",full.names=T,recursive = T)
dat<-bind_rows(lapply(f,read.csv))

dat$timestamp<-as.POSIXct(dat$End,format="%H:%M:%S %d-%b-%Y",tz="GMT")

dat$Month<-months(dat$timestamp)
dat$Month<-factor(dat$Month,levels=month.name)
dat$Hour<-strftime(dat$timestamp,format="%H")
dat$Year<-years(dat$timestamp)

### for testing
gdat<-gdat %>% filter(Ptt %in% unique(dat$Ptt))  

gdat<-gdat %>% dplyr::select(Animal=Ptt,timestamp,Quality,Latitude,Longitude)
gdat$Month<-months(gdat$timestamp)
gdat$Month<-factor(gdat$Month,levels=month.name)

dive<-dat %>% filter(What=="Dive")%>% dplyr::select(Animal=Ptt,timestamp,Hour,Month,Year,DepthMax,DepthMin,DurationMax,DurationMin)
dive<-bind_rows(gdat,dive)

#order by timestamp
dive<-dive %>% arrange(timestamp)
mdat<-dive
```

## Dive profiles per indidivuals

```{r,fig.height=11,eval=T}
ggplot(dive[,],aes(x=timestamp,y=DepthMax)) + geom_point(size=0.1) + geom_line(size=0.1) + facet_wrap(~Animal,scales="free",ncol=2) + theme_bw() + labs(x="Date",y="Dive Depth (m)") + scale_y_reverse()
ggsave("Figures/perindividual.jpeg",height=12,width=13)
ggsave("Figures/perindividual.svg",height=12,width=13)
```

Dive Profiles with Argos timestamps

```{r,fig.height=11,eval=F}
ggplot(dive[,],aes(x=timestamp,y=-DepthMax)) + geom_point(size=0.1) + geom_line(size=0.1) + facet_wrap(~Animal,scales="free",ncol=2) + theme_bw() + geom_point(data=dive[!is.na(dive$Latitude),],y=0,col="red",size=1,aes(x=timestamp))
ggsave("Figures/perindividual.jpeg",height=6,width=12)
```

Dive autocorrelation plots

```{r,eval=T}
acf_df<-function(x){
  bacf <- acf(x, plot = FALSE)
  bacfdf <- with(bacf, data.frame(lag, acf))
}
acplot<-dive %>% group_by(Animal) %>% arrange(timestamp) %>% filter(!is.na(DepthMax)) %>% do(acf_df(.$DepthMax))
q <- ggplot(data = acplot, mapping = aes(x = lag, y = acf)) +
       geom_hline(aes(yintercept = 0)) +
       geom_segment(mapping = aes(xend = lag, yend = 0))
q + facet_wrap(~Animal)
```

```{r,eval=T}
mybbox<-make_bbox(data=mdat,lat=Latitude,lon=Longitude,f=0.1)
troy <- get_map(location = mybbox, maptype = "toner-background")

attr_troy <- attr(troy, "bb")    # save attributes from original

# change color in raster
troy[troy == "#FFFFFF"] <- "#C0C0C0"
troy[troy == "#000000"] <- "#FFFFFF"

# correct class, attributes
class(troy) <- c("ggmap", "raster")
attr(troy, "bb") <- attr_troy

#nice to have it as a function
makemap<-function(x){
  mybbox<-make_bbox(data=x,lat=Latitude,lon=Longitude,f=0.1)
troy <- get_map(location = mybbox, maptype = "toner-background")

attr_troy <- attr(troy, "bb")    # save attributes from original

# change color in raster
troy[troy == "#FFFFFF"] <- "#C0C0C0"
troy[troy == "#000000"] <- "#FFFFFF"

# correct class, attributes
class(troy) <- c("ggmap", "raster")
attr(troy, "bb") <- attr_troy
return(troy)
}
```

## Data Statistics before track cut

```{r}
mdat %>% group_by(Animal) %>% summarize(n=n(),argos=sum(!is.na(Latitude)),dive=sum(!is.na(DepthMax)))
nrow(mdat)
```

```{r,fig.width=12,fig.height=10}
#set max depth to km
#mdat$DepthMax<-mdat$DepthMax/1000

#Specifiy local time
  
#local time
mdat$localtime<-as.POSIXct(format(mdat$timestamp,tz="etc/GMT+3"))
mdat$LocalHour<-as.numeric(strftime(mdat$localtime,format="%H"))

#set duration to minutes
mdat$DurationMax<-mdat$DurationMax/60

#view data
ggmap(troy) + geom_point(data=mdat,aes(x=Longitude, y=Latitude,col=as.factor(Animal)),size=0.3)  + mytheme + labs(col="Animal")
ggsave("Figures/Map.svg")
ggsave("Figures/Map.png",height=6,width=8)

```

```{r,eval=F}
#By individual

for(x in unique(mdat$Animal)){
  mmap<-makemap(mdat %>% filter(Animal==x))
  ggmap(mmap)+geom_path(data=mdat,aes(x=Longitude, y=Latitude,group=Animal),size=.5) + geom_point(data=mdat,aes(x=Longitude, y=Latitude,col=DepthMax)) + theme_bw() + mytheme + scale_color_continuous(low="blue",high="red") + labs(col="Max Dive Depth (km)") + #facet_wrap(~Animal,ncol=2)
}

```

```{r,eval=F}
server <- function(input, output) {

# create a reactive value that will store the click position
data_of_click <- reactiveValues(clickedMarker=NULL)

dat<-mdat %>% filter(Animal=="131130",!is.na(Latitude)) %>% mutate(ID=1:nrow(.))

# Leaflet map with 2 markers
pal<-colorFactor(heat.colors(10),dat$DepthMax)

output$map <- renderLeaflet({
leaflet() %>% 
 setView(lng=mean(dat$Longitude,na.rm=T), lat =mean(dat$Latitude,na.rm=T), zoom=8) %>%
 addTiles(options = providerTileOptions(noWrap = TRUE)) %>% addPolylines(data=dat %>% arrange(Animal,timestamp),~Longitude, ~Latitude,weight=0.25) %>%
 addCircleMarkers(data=dat, ~Longitude , ~Latitude, layerId=~ID, popup=~paste(timestamp), radius=3 , fillColor="red", stroke = TRUE, fillOpacity = 0.1) 
})

# store the click
observeEvent(input$map_marker_click,{
data_of_click$clickedMarker <- input$map_marker_click
})

# Make a barplot or scatterplot depending of the selected point
output$plot=renderPlot({
my_place=dat[dat$ID %in% data_of_click$clickedMarker$id,] 
print(data_of_click$clickedMarker)
print(my_place)

p<-ggplot(mdat[mdat$Animal %in% dat$Animal,],aes(x=timestamp,y=-DepthMax)) + geom_point() + geom_line(size=0.5) + theme_bw() 
if(!is.null(my_place)){
p<-p + geom_vline(data=my_place,aes(xintercept=timestamp),size=4,col='red',alpha=0.5) 
}
return(p)
})
}


ui <- fluidPage(
br(),
leafletOutput("map", height="600px"),
plotOutput("plot", height="300px"),
br()
)

shinyApp(ui = ui, server = server)
```

#Data Cleaning Rules.

1. All tracks start with an observed argos location.
2. A single track must have one argos location per time step.

```{r}

##Time is the beginning of the first point.
step_length=6

sxy<-split(mdat,mdat$Animal)

#Cut into tracks
#time diff function
timed<-function(d,step_length){
  
  #Order and startmwith a valid observation
  d<-d %>% arrange(timestamp)
  startpoint<-min(which(!is.na(d$Latitude)))
  d<-d[startpoint:nrow(d),]
  d$interval <- cut(d$timestamp,breaks = seq(min(d$timestamp), max(d$timestamp)+step_length*3600, as.difftime(step_length, units="hours")))
  
  #If there are no argos observations, remove interval.
  remove_interval<-d %>% group_by(interval) %>% summarize(n=sum(!is.na(Latitude))) %>% filter(n==0)
  d<-d %>% filter(!interval %in% remove_interval$interval )
  
  #No empty dive intervals
  remove_dive<-d %>% group_by(interval) %>% summarize(n=sum(!is.na(DepthMax))) %>% filter(n==0)
  d<-d %>% filter(!interval %in% remove_dive$interval )
  
  #Can we have empty intervals?
  d<-d %>% filter(!is.na(interval)) %>% droplevels()
  
  #which intervals are more than 12 hours apart
  Track=c()
  Track[1]<-1
  counter=1
  
  for(x in 2:length(unique(d$interval))){
    difft<-as.numeric(difftime(as.POSIXct(unique(d$interval)[x]),as.POSIXct(unique(d$interval)[x-1]),units="hours"))
    if(difft>step_length){
      counter=counter+1
    } 
    Track[x]<-counter
  }
  
  d<-d %>% inner_join(data.frame(interval=unique(d$interval),Track))
  
  #First position in each track must be a valid position, recode 
  d<-d %>% group_by(Track) %>% filter(timestamp >= min(timestamp[!is.na(Latitude)]))
  
  #get jStep and interval time
  d<-d %>% group_by(interval) %>% mutate(j=difftime(timestamp,as.POSIXct(interval,tz="GMT"),units="hours")/step_length,jStep=1:n())
  
  #set step and remove tracks less than 2 steps, a bit ugly to maintain subgroup order.
  d<-d %>% group_by(Track) %>% mutate(step=as.numeric(as.factor(as.character(interval)))) %>% filter(max(step) > 2)
  
  #reset track numbering
  #need more than 3 jSteps
    #d<-d %>% group_by(Track,step) %>% filter(max(jStep) > 3)

  #remove tracks that are shorter than 12 hours
  track_time<-d %>% group_by(Track) %>% summarize(mt=difftime(max(as.POSIXct(timestamp)),min(as.POSIXct(timestamp)),units="hours")) %>% filter(mt>=12) %>% .$Track
  
  return(d)
  }

sxy<-lapply(sxy,timed,step_length=step_length)
mdat<-bind_rows(sxy)

#recode tracks and steps to make ordinal
mdat<-mdat %>% group_by(Animal) %>% mutate(Track=as.numeric(as.factor(Track)))
```

## Data after track cut
```{r,fig.height=12,fig.width=12}
mdat %>% group_by(Animal) %>% summarize(n=n())
mdat %>% group_by(Animal) %>% summarize(Tracks=length(unique(Track)))
ggplot(mdat[,],aes(x=timestamp,y=-DepthMax)) + geom_point(size=0.1) + geom_line(size=0.1) + facet_wrap(~Animal,scales="free",ncol=2) + theme_bw() + geom_point(data=dive[!is.na(dive$Latitude),],y=0,col="red",size=1,aes(x=timestamp))

```

```{r}
###################################################
#filter by two whales for the moment
mdat<-mdat %>% filter(!Animal %in% c("131132","131127"))
#mdat<-mdat %>% filter(Animal %in% "131115")

#remake map
troy<-makemap(mdat)
####################################################

#refactor animal
mdat$jAnimal<-as.numeric(as.factor(mdat$Animal))

##Split into format for JAGS
#Cast time array
j<-reshape2::acast(mdat,jAnimal~Track~step~jStep,value.var="j")

#how many observations per individual in each step
idx<-mdat %>% group_by(jAnimal,Track,step) %>% summarize(n=n())
colnames(idx)<-c("Animal","Track","step","jStep")
idx<-reshape2::acast(data=idx,Animal~Track~step)

#Individuals
ind=length(unique(mdat$Animal))

#tracks per indivudal
tracks<-mdat %>% group_by(Animal) %>% summarize(tracks=length(unique(Track))) %>% .$tracks

#steps per track
steps_all<-mdat %>% group_by(Animal,Track) %>% summarize(n=length(unique(step)))
steps<-reshape2::acast(steps_all,Animal~Track,value.var="n")

#obs array
obs<-reshape2::melt(mdat,measure.vars=c("Longitude","Latitude"))
obs<-reshape2::acast(obs,Animal~Track~step~jStep~variable,fun.aggregate = mean)
obs[!is.finite(obs)]<-NA

#argos class array
mdat$argos.lc<-factor(mdat$Quality,levels=c(3,2,1,0,"A","B"))
mdat$numargos<-as.numeric(mdat$argos.lc)
obs_class<-reshape2::acast(mdat,Animal~Track~step~jStep,value.var="numargos",fun.aggregate = min)
#set interpolated observations to having lowest class of argos error
obs_class[!is.finite(obs_class)]<-6

#average dive depth array
maxdive<-reshape2::acast(mdat,Animal~Track~step~jStep,value.var="DepthMax",fun.aggregate = mean)

hours<-reshape2::acast(mdat,Animal~Track~step~jStep,value.var="LocalHour")

#fill the empty values
maxdive[!is.finite(maxdive)]<-NA

```

```{r,child="Bayesian/NestedDive.R",eval=T}
```

```{r,eval=T}
#source jags file
source("Bayesian/NestedDive.R")

#prior cov shape
R <- diag(c(1,1))
data=list(hours=hours,divedepth=maxdive,argos=obs,steps=steps,R=R,ind=ind,j=j,idx=idx,tracks=tracks,argos_class=obs_class)

#Inits
stage_init<-hours
stage_init[]<-1

InitStage <- function() {list(stage=array(data=2,dim=c(dim(hours))))}

#paramters to track
pt<-c("alpha","sub_alpha","gamma","depth_mu","depth_tau","dive_new","state", "sub_state","E","Enew")

if(newModel){
  system.time(diving<-jags(model.file = "Bayesian/NestedDive.jags",data=data,n.chains=2,parameters.to.save=pt,n.iter=10000,n.burnin=9500,n.thin=2,DIC=FALSE,inits = InitStage))
  }

```

##Chains

```{r,eval=T}
#bind chains
pc_dive<-reshape2::melt(diving$BUGSoutput$sims.array)
colnames(pc_dive)<-c("Draw","chain","par","value")

#extract parameter name
pc_dive$parameter<-data.frame(str_match(pc_dive$par,"(\\w+)"))[,-1]

#Extract index
splitpc<-split(pc_dive,pc_dive$parameter)

#single index
splitpc[c("alpha","gamma")]<-lapply(splitpc[c("alpha","gamma")],function(x){
    sv<-data.frame(str_match(x$par,"(\\w+)\\[(\\d+)]"))[,3]
    pc<-data.frame(x,Behavior=sv)
    return(pc)
}) 

#double index
splitpc[c("depth_mu","depth_tau","sub_alpha")]<-lapply(splitpc[c("depth_mu","depth_tau","sub_alpha")],function(x){
    sv<-data.frame(str_match(x$par,"(\\w+)\\[(\\d+),(\\d+)"))[,3:4]
    colnames(sv)<-c("Behavior","sub_state")
    pc<-data.frame(x,sv)
    return(pc)
}) 

#3 index
splitpc[c("state")]<-lapply(splitpc[c("state")],function(x){
    sv<-data.frame(str_match(x$par,"(\\w+)\\[(\\d+),(\\d+),(\\d+)"))[,3:5]
    colnames(sv)<-c("jAnimal","Track","step")
    pc<-data.frame(x,sv)
    return(pc)
}) 

#4 index
splitpc[c("dive_new","E","Enew","sub_state")]<-lapply(splitpc[c("dive_new","E","Enew","sub_state")],function(x){
    sv<-data.frame(str_match(x$par,"(\\w+)\\[(\\d+),(\\d+),(\\d+),(\\d+)]"))[,3:6]
    colnames(sv)<-c("jAnimal","Track","step","jStep")
    pc<-data.frame(x,sv)
    return(pc)
}) 

#bind all matrices back together
pc_dive<-bind_rows(splitpc)
rm(splitpc)

ggplot(pc_dive[!pc_dive$parameter %in% c("dive_new","state","E","Enew","eval","sub_state"),],aes(x=Draw,y=value,col=as.factor(chain))) + geom_line() + facet_wrap(~par,scales="free")

#posteriors
ggplot(pc_dive[!pc_dive$parameter %in% c("dive_new","state","E","Enew","eval","sub_state"),],aes(x=value,fill=as.factor(chain))) + geom_histogram() + facet_wrap(~par,scales="free")

#sum table
pc_dive %>% filter(!parameter %in% c("dive_new","state","E","Enew","eval","sub_state")) %>% group_by(parameter,Behavior,sub_state) %>% summarize(mean=round(mean(value),3),upper=round(quantile(value,0.95),3),lower=round(quantile(value,0.05),3)) 

#Animal lookup table
pc_dive<-mdat %>% ungroup() %>%  dplyr::select(Animal,jAnimal)  %>% distinct() %>% mutate(jAnimal=as.factor(jAnimal)) %>% full_join(pc_dive)
```

# Temporal Variation in Dive Behavior

```{r}

#Take the most common estimate of behavior
Mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

#Combine posterior summary
mdat$Animal<-as.factor(mdat$Animal)
mdat$Track<-as.factor(mdat$Track)
mdat$step<-as.factor(mdat$step)

state_est<-pc_dive %>% group_by(jAnimal,Track,step,parameter) %>%  filter(parameter %in% c("state","sub_state")) %>% summarize(Behavior=Mode(value)) %>% ungroup()  %>% mutate(jAnimal=as.numeric(as.character(jAnimal)),Track=as.numeric(as.character(Track)),step=as.numeric(as.character(step))) %>% spread(parameter,Behavior) 

state_est %>% group_by(state,sub_state) %>% summarize(n=n())
state_est<-state_est %>% inner_join(mdat)

blookup<-data.frame(state=c(1,1,2,2),sub_state=c(1,2,1,2),Behavior=c("Traveling","Dummy","Foraging","Resting"))
state_est<-state_est %>% inner_join(blookup)
```

```{r,fig.height=13,fig.width=20}
ggplot(state_est %>% filter(!is.na(DepthMax)),aes(x=timestamp,y=DepthMax,col=Behavior,group=Track)) + geom_point() + geom_line(size=0.1,aes(group=Track)) + facet_wrap(~Animal,scales="free",ncol=1) + theme_bw() + scale_y_reverse() + labs(x="Date",y="Dive Depth (m)")
ggsave("Figures/TemporalBehavior.svg",height=12,width=10)
ggsave("Figures/TemporalBehavior.jpeg",height=12,width=10)
```

Grouped by stage

```{r}
ggplot(state_est %>% filter(!is.na(DepthMax)),aes(x=timestamp,y=-DepthMax,col=as.factor(state),group=step)) + geom_point() + geom_line(size=0.25,aes(group=Track)) + facet_wrap(~Animal,scales="free",ncol=1) + theme_bw()
```

# Behavior Sequence
```{r}
ggplot(state_est,aes(x=timestamp,y=Behavior)) + geom_line(aes(group=1))  + geom_point() + facet_wrap(~Animal,scales="free",ncol=1) + theme_bw()

```

## Simulate dive depths

```{r}
#how many of each state to draw?
statecount<-state_est %>% group_by(Behavior) %>% filter(is.na(Latitude)) %>% summarize(n=length(jStep))

pred_dives<-list()
for(x in 1:nrow(statecount)){
  state_index<-blookup[ blookup$Behavior %in% statecount$Behavior[x],"state"]
  sub_state_index<-blookup[ blookup$Behavior %in%  statecount$Behavior[x],"sub_state"]

  depth_tau<-pc_dive %>% filter(parameter %in% c("depth_tau"),Behavior==state_index,sub_state==sub_state_index)
  depth_mu<-pc_dive %>% filter(parameter %in% c("depth_mu"),Behavior==state_index,sub_state==sub_state_index)

  travel_dives<-data.frame(DepthMax=rtruncnorm(n=as.numeric(statecount[x,"n"]),mean=depth_mu$value,sd=1/sqrt(depth_tau$value),a=0),Behavior=statecount[x,"Behavior"])
  pred_dives[[x]]<-travel_dives
}

pred_dives<-bind_rows(pred_dives)

ggplot(pred_dives) + geom_histogram(alpha=0.8,aes(x=DepthMax,fill=Behavior))  + theme_bw() + labs(x="Dive Depth (m)") +facet_wrap(~Behavior,scales="free")

ggplot(pred_dives)  + geom_histogram(data=mdat,aes(x=DepthMax),col="black")+ geom_histogram(alpha=0.8,aes(x=DepthMax,fill=Behavior))  + theme_bw() + labs(x="Dive Depth (m)")
ggsave("Figures/DiveHist.jpg",height=4,width=7)


ggplot(pc_dive[pc_dive$parameter=="dive_new",])+  geom_density(aes(x=value),col='red') + geom_density(data=mdat,aes(x=DepthMax),col="black",size=1.5) + theme_bw() + labs(x="Dive Depth (m)") 
ggsave("Figures/DivePredict.jpg",height=4,width=7) 

byanimal<-pc_dive[pc_dive$parameter=="dive_new",]
ggplot(mdat)+  geom_density(aes(x=DepthMax),col="black",size=1.5) +  facet_wrap(~Animal,scales="free") + geom_density(data=byanimal,aes(x=value),col='red') + theme_bw() + labs(x="Dive Depth (m)")  

```

## Diel

```{r}
ggplot(data=state_est) + geom_boxplot(aes(x=as.factor(LocalHour),y=DepthMax,fill=Behavior)) + ylab("Dive Depth (m)")+ labs(x="Hour (Local GMT+3)")
ggsave("Figures/Diel.svg")
ggsave("Figures/Diel.png",height=5,width=8,unit="in")

ggplot(data=state_est) + geom_boxplot(aes(x=as.factor(LocalHour),y=DepthMax,fill=Behavior)) + ylab("Dive Depth (m)") + facet_wrap(~Month) + labs(x="Hour (Local GMT+3)")
ggsave("Figures/DielbyMonth.svg")
ggsave("Figures/DielbyMonth.png",height=5,width=10,unit="in")
```

## Month
```{r}
ggplot(data=state_est) + geom_boxplot(aes(x=Month,y=DepthMax,fill=Behavior)) + ylab("Maximum Dive Depth (km)")
ggsave("Figures/Month.svg")
ggsave("Figures/Month.png",height=5,width=8,unit="in")
```

# Spatial Prediction

```{r,fig.height=10,fig.width=10}
state_est<-state_est %>% arrange(Animal,Track,timestamp)
ggmap(troy) +geom_path(data=state_est %>% filter(!is.na(Latitude)),aes(x=Longitude, y=Latitude,group=paste(Animal,Track)),size=0.5) + geom_point(data=state_est,aes(x=Longitude, y=Latitude,col=Behavior),size=0.5) + mytheme 

ggmap(troy) +geom_path(data=state_est %>% filter(!is.na(Latitude)),aes(x=Longitude, y=Latitude,group=paste(Animal,Track)),size=0.5) + geom_point(data=state_est,aes(x=Longitude, y=Latitude,col=Behavior),size=0.5) + mytheme +facet_wrap(~Animal)

#Show nested behaviors
state_est$TopLayer<-NA
state_est[state_est$state==1,"TopLayer"]<-"Traveling"
state_est[state_est$state==2,"TopLayer"]<-"Area-restricted Search"
state_est$SecondLayer<-state_est$Behavior
state_est<-gather(state_est,key="Layer",value="estimate",TopLayer:SecondLayer)

state_est$Layer<-factor(state_est$Layer,levels = c("TopLayer","SecondLayer"),labels=c("Top Layer","Second Layer"))
ggmap(troy) +geom_path(data=state_est %>% filter(!is.na(Latitude)),aes(x=Longitude, y=Latitude,group=paste(Animal,Track)),size=0.5) + geom_point(data=state_est,aes(x=Longitude, y=Latitude,col=estimate),size=0.5) + mytheme +facet_wrap(~Layer) + scale_color_manual(values=c("plum","Red","Green","Blue")) + labs(col="Behavior")
ggsave("Figures/NestedMap.jpeg",height=8,width=11)
```

#Residual Explorations
##Diel

```{r}
ggplot(state_est,aes(x=LocalHour,y=DepthMax,col=Behavior)) + geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) + facet_wrap(~Animal)
```

```{r,fig.height=12}
byhour<-state_est %>% group_by(Animal,LocalHour,Behavior) %>% summarize(n=n()) %>% mutate(prop=n/sum(n))
ggplot(byhour,aes(x=LocalHour,y=prop,col=Behavior)) + geom_line()  + geom_point() + facet_wrap(~Animal,scales="free",ncol=1) + theme_bw() + scale_y_continuous(labels=scales::percent)
```

Lines connect individuals
```{r}
byhour<-state_est %>% group_by(LocalHour,Behavior) %>% summarize(n=n()) %>% mutate(prop=n/sum(n))
ggplot(byhour,aes(x=LocalHour,y=prop,col=Behavior)) + geom_line()  + geom_point() + theme_bw() + scale_y_continuous("Frequency of Behavior",labels=scales::percent) 
```

## Monthly

Lines connect individuals
```{r}
bymonth<-state_est %>% group_by(Animal,Month,Behavior) %>% summarize(n=n()) %>% mutate(prop=n/sum(n)) %>% filter(!is.na(Month)) %>% filter(Behavior %in% c("Foraging","Resting"))
bymonth$Month<-factor(bymonth$Month,levels=month.name)
ggplot(bymonth,aes(x=Month,y=prop,col=Behavior)) + geom_line(aes(group=paste(Animal,Behavior)))  + geom_point() + theme_bw() + scale_y_continuous("Frequency of Behavior",labels=scales::percent)
ggplot(bymonth,aes(x=Month,y=prop,fill=Behavior)) + geom_boxplot() + theme_bw() + scale_y_continuous("Frequency of Behavior",labels=scales::percent)
ggsave("Figures/MonthFreq.svg",height=6,width=9)
ggsave("Figures/MonthFreq.png",height=6,width=9)
```

## Posterior Checks

The goodness of fit is a measured as chi-squared. The expected value is compared to the observed value of the actual data. In addition, a replicate dataset is generated from the posterior predicted intensity. Better fitting models will have lower discrepancy values and be 
Better fitting models are smaller values and closer to the 1:1 line. A perfect model would be 0 discrepancy. This is unrealsitic given the stochasticity in the sampling processes. Rather, its better to focus on relative discrepancy. In addition, a model with 0 discrepancy would likely be seriously overfit and have little to no predictive power.

```{r,fig.height=4,fig.width=8,message=F,warning=F}

fitstat<-pc_dive %>% filter(parameter %in% c("E","Enew")) %>% group_by(parameter,Draw,chain) %>% summarize(fit=mean(value))

fitstat<-dcast(fitstat,Draw+chain~parameter,value.var="fit")

ymin<-min(c(fitstat$E,fitstat$Enew)) - min(c(fitstat$E,fitstat$Enew)) * .1
ymax<-max(c(fitstat$E,fitstat$Enew)) + max(c(fitstat$E,fitstat$Enew)) * .1
ggplot(fitstat,aes(x=E,y=Enew)) + geom_point() + theme_bw() + labs(x="Discrepancy of observed data",y="Discrepancy of replicated data") + geom_abline() + coord_fixed() + ylim(ymin=ymin,ymax=ymax) + xlim(xmin=ymin,xmax=ymax)


ggsave("Figures/PosteriorCheck.jpg",height = 4,width = 4)
fitstat %>% group_by() %>% summarize(mean(E),var(Enew))

#By Animal

fitstat<-pc_dive %>% filter(parameter %in% c("E","Enew")) %>% group_by(parameter,Draw,chain,jAnimal) %>% summarize(fit=mean(value))

fitstat<-dcast(fitstat,jAnimal+Draw+chain~parameter,value.var="fit")

ymin<-min(c(fitstat$E,fitstat$Enew)) - min(c(fitstat$E,fitstat$Enew)) * .1
ymax<-max(c(fitstat$E,fitstat$Enew)) + max(c(fitstat$E,fitstat$Enew)) * .1
ggplot(fitstat,aes(x=E,y=Enew)) + geom_point() + theme_bw() + labs(x="Discrepancy of observed data",y="Discrepancy of replicated data") + geom_abline() + coord_fixed() + facet_wrap(~jAnimal,scale="free")

ggsave("Figures/PosteriorCheckAnimal.jpg",height = 4,width = 4)
fitstat %>% group_by() %>% summarize(mean(E),var(Enew))
```

```{r}
#save.image("/Users/Ben/Dropbox/Whales/Dive/WhalePhys.RData")
```