---
title: "Foraging Model"
author: "Ben Weinstein"
date: "`r Sys.time()`"
output: 
  html_document:
    toc: true
    number_sections: true
    theme: united
---


# Foraging Model

The proportion of time an animal is in a feeding behavioral state.

(@eq1)

*Process Model*

$$Y_{i,t+1} \sim Multivariate Normal(d_{i,t},σ)$$

$$d_{i,t}= Y_{i,t} + γ_{s_{i,g,t}}*T_{i,g,t}*( Y_{i,g,t}- Y_{i,g,t-1} )$$

$$
\begin{matrix}
  \alpha_{i,1,1} & \beta_{i,1,1} & 1-(\alpha_{i,1,1} + \beta_{i,1,1}) \\
  \alpha_{i,2,1} & \beta_{i,2,1} & 1-(\alpha_{i,2,1} + \beta_{i,2,1}) \\
    \alpha_{i,3,1} & \beta_{i,3,1} & 1-(\alpha_{i,3,1} + \beta_{i,3,1}) \\
\end{matrix}
$$
$$logit(\phi_{Behavior}) = \alpha_{Behavior_{t-1}} $$
The behavior at time t of individual i on track g is a discrete draw.
$$S_{i,g,t} \sim Cat(\phi_{traveling},\phi_{foraging},\phi_{resting})$$

Dive information is a mixture model based on behavior (S)

$\text{Average dive depth}(\psi)$
$$ DiveDepth \sim Normal(dive_{\mu_S},dive_{\tau_S})$$
$$ DiveDuration \sim Normal(duration_{\mu_S},duration_{\tau_S})$$


```{r,echo=F,warning=F,message=F}
library(tidyr)
library(ggplot2)
library(maptools)
library(shiny)
library(raster)
library(data.table)
library(ggmap)
library(leaflet)
library(dplyr)
library(stringr)
library(chron)
library(R2jags)
library(boot)
library(knitr)
library(MCMCpack)
library(truncnorm)

newModel=T

opts_chunk$set(echo=F,warning=F,message=F,fig.width=11)

mytheme<-theme(axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.text.y=element_blank(),axis.ticks.y=element_blank(),axis.title.x=element_blank(),axis.title.y=element_blank(),panel.grid=element_blank())
```

```{r}
if(!newModel){
  load("/Users/Ben/Dropbox/Whales/Dive/WhalePhys.RData")
  newModel=F
}

#get gps data
f<-list.files("/Users/Ben/Dropbox/Whales/Data/Humpback",pattern="Locations",full.names=T,recursive = T)
gdat<-lapply(f,function(x) read.csv(x,stringsAsFactors=F))
gdat<-lapply(gdat,function(x){
  x$Quality<-as.character(x$Quality)
  return(x)
}) 

gdat<-bind_rows(gdat)

#timestamp
gdat$timestamp<-as.POSIXct(gdat$Date,format="%H:%M:%S %d-%b-%Y",tz="GMT")

#
gdat<-gdat[!is.na(gdat$Latitude),]

#crop by extent
d<-SpatialPointsDataFrame(cbind(gdat$Longitude,gdat$Latitude),data=data.frame(gdat),proj4string=CRS("+proj=longlat +datum=WGS84"))

cropoly<-readShapePoly("Data/CutPolygon.shp",proj4string=CRS("+proj=longlat +datum=WGS84"))

b<-d[!is.na(d %over% cropoly)[,2],]

gdat<-b@data

#get dive data files
f<-list.files("/Users/Ben/Dropbox/Whales/Data/Humpback/",pattern="Behavior",full.names=T,recursive = T)
dat<-bind_rows(lapply(f,read.csv))

dat$timestamp<-as.POSIXct(dat$End,format="%H:%M:%S %d-%b-%Y",tz="GMT")

dat$Month<-months(dat$timestamp)
dat$Month<-factor(dat$Month,levels=month.name)
dat$Hour<-strftime(dat$timestamp,format="%H")
dat$Year<-years(dat$timestamp)

### for testing
#dat<-dat %>% filter(Ptt=="131134")
gdat<-gdat %>% filter(Ptt %in% unique(dat$Ptt))  

gdat<-gdat %>% dplyr::select(Animal=Ptt,timestamp,Quality,Latitude,Longitude)
dive<-dat %>% filter(What=="Dive")%>% dplyr::select(Animal=Ptt,timestamp,Hour,Month,Year,DepthMax,DepthMin,DurationMax,DurationMin)
dive<-bind_rows(gdat,dive)

#order by timestamp
dive<-dive %>% arrange(timestamp)
```

## Dive profiles per indidivuals

```{r,fig.height=12}
ggplot(dive[,],aes(x=timestamp,y=-DepthMax)) + geom_point(size=0.1) + geom_line(size=0.1) + facet_wrap(~Animal,scales="free",ncol=3) + theme_bw()
ggsave("Figures/perindividual.jpeg",height=6,width=12)
```

Dive Profiles with Argos timestamps

```{r}
ggplot(dive[,],aes(x=timestamp,y=-DepthMax)) + geom_point(size=0.1) + geom_line(size=0.1) + facet_wrap(~Animal,scales="free",ncol=2) + theme_bw() + geom_point(data=dive[!is.na(dive$Latitude),],y=0,col="red",size=1,aes(x=timestamp))
ggsave("Figures/perindividual.jpeg",height=6,width=12)
```

```{r}
dive %>% group_by(Animal,Year) %>% summarize(n=n())
```

```{r}
mdat<-dive

mybbox<-make_bbox(data=mdat,lat=Latitude,lon=Longitude,f=0.25)
troy <- get_map(location = mybbox, maptype = "toner-background")

attr_troy <- attr(troy, "bb")    # save attributes from original

# change color in raster
troy[troy == "#FFFFFF"] <- "#C0C0C0"
troy[troy == "#000000"] <- "#FFFFFF"

# correct class, attributes
class(troy) <- c("ggmap", "raster")
attr(troy, "bb") <- attr_troy

#nice to have it as a function
makemap<-function(x){
  mybbox<-make_bbox(data=x,lat=Latitude,lon=Longitude,f=0.25)
troy <- get_map(location = mybbox, maptype = "toner-background")

attr_troy <- attr(troy, "bb")    # save attributes from original

# change color in raster
troy[troy == "#FFFFFF"] <- "#C0C0C0"
troy[troy == "#000000"] <- "#FFFFFF"

# correct class, attributes
class(troy) <- c("ggmap", "raster")
attr(troy, "bb") <- attr_troy
return(troy)
}
```

## Data Statistics before track cut

```{r}
mdat %>% group_by(Animal) %>% summarize(n=n())
nrow(mdat)
```


```{r}
#set max depth to km
#mdat$DepthMax<-mdat$DepthMax/1000

#Specifiy local time
  
#local time
mdat$localtime<-as.POSIXct(format(mdat$timestamp,tz="etc/GMT+3"))
mdat$LocalHour<-strftime(mdat$localtime,format="%H")

#set duration to minutes
mdat$DurationMax<-mdat$DurationMax/60

#dives can't be longer than 20 minutes
mdat<-mdat %>% filter(DurationMax<20)

#view data
ggmap(troy)+geom_path(data=mdat,aes(x=Longitude, y=Latitude,group=Animal),size=.5) + geom_point(data=mdat,aes(x=Longitude, y=Latitude)) + theme_bw() + mytheme + scale_color_continuous(low="blue",high="red") + labs(col="Max Dive Depth (km)")
#ggsave("Figures/Map.svg")
#ggsave("Figures/Map.png")
```

```{r,eval=F}
#By individual

#for(x in unique(mdat$Animal)){
#  mmap<-makemap(mdat %>% filter(Animal==x))
#  ggmap(mmap)+geom_path(data=mdat,aes(x=Longitude, y=Latitude,group=Animal),size=.5) + geom_point(data=mdat,aes(x=Longitude, # y=Latitude,col=DepthMax)) + theme_bw() + mytheme + scale_color_continuous(low="blue",high="red") + labs(col="Max Dive Depth (km)") + #facet_wrap(~Animal,ncol=2)
}

```

```{r,eval=F}
server <- function(input, output) {

# create a reactive value that will store the click position
data_of_click <- reactiveValues(clickedMarker=NULL)

dat<-mdat %>% filter(Animal=="131134")

# Leaflet map with 2 markers
pal<-colorFactor(heat.colors(10),dat$DepthMax)

output$map <- renderLeaflet({
leaflet() %>% 
 setView(lng=mean(dat$Longitude), lat =mean(dat$Latitude), zoom=8) %>%
 addTiles(options = providerTileOptions(noWrap = TRUE)) %>% addPolylines(data=dat %>% arrange(Animal,timestamp),~Longitude, ~Latitude,weight=0.25) %>%
 addCircleMarkers(data=dat, ~Longitude , ~Latitude, layerId=~ID, popup=~paste(timestamp,DepthMax,DurationMax), radius=2 ,color=~pal(DepthMax), fillColor="red", stroke = TRUE, fillOpacity = 0.1) 
})

# store the click
observeEvent(input$map_marker_click,{
data_of_click$clickedMarker <- input$map_marker_click
})

# Make a barplot or scatterplot depending of the selected point
output$plot=renderPlot({
my_place=data_of_click$clickedMarker$id

if(!is.null(my_place)){
  animal_selected<-mdat %>% filter(my_place==ID) %>% .$Animal
ggplot(mdat[mdat$Animal == animal_selected,],aes(x=timestamp,y=-DepthMax)) + geom_point(aes(col=my_place==ID,size=my_place==ID)) + geom_line(size=0.1) + theme_bw() + scale_color_manual(values=c("Black","Red"),guide=F) + scale_size_manual(values=c(0.1,5),guide=F)
}

})
}


ui <- fluidPage(
br(),
leafletOutput("map", height="600px"),
plotOutput("plot", height="300px"),
br()
)

shinyApp(ui = ui, server = server)
```

#Data Cleaning Rules.

1. All tracks start with an observed argos location.
2. A single track must have one argos location per time step.

```{r}

##Time is the beginning of the first point.
step_length=6

sxy<-split(mdat,mdat$Animal)

d<-d %>% arrange(timestamp)
startpoint<-which.min(!is.na(d$Latitude))
d<-d[startpoint:nrow(d),]

d$interval <- cut(d$timestamp,breaks = seq(min(d$timestamp), max(d$timestamp), as.difftime(step_length, units="hours")))

as.xts(d$timestamp)

#cut time into length steps
cut(d$timestamp,6)

#time diff function
timed<-function(d,step_length){
  
  d$j[1]<-0
  for (x in 2:nrow(d)){
    d$j[x]<-as.numeric(difftime(as.POSIXct(d$timestamp[x]),as.POSIXct(d$timestamp[x-1]),units="mins"))/(step_length*60)
  }
  
  #Split out track endings
  ends<-c(1,which(d$j>1),nrow(d))

  for(w in 2:length(ends)){
    d[ends[w-1]:ends[w],"Track"]<-w-1
  }
  
  #remove tracks that are shorter than 1 days
  track_time<-d %>% group_by(Track) %>% summarize(mt=difftime(max(as.POSIXct(timestamp)),min(as.POSIXct(timestamp)),units="hours")) %>% filter(mt>=24) %>% .$Track
  
  d<-d[d$Track %in% track_time,]
  
  #renumber the tracks
  d$Track<-as.numeric(as.factor(d$Track))
  return(d)
  }

sxy<-lapply(sxy,timed,step_length=step_length)

#Format matrices for jags
mdat<-bind_rows(sxy)

sxy<-split(mdat,list(mdat$Animal,mdat$Track),drop=TRUE)

sxy<-lapply(sxy,function(x){
#How many observations in each step length segment
x$step<-as.numeric(cut(as.POSIXct(x$timestamp),paste(step_length,"hours")))
return(x)
})

mdat<-bind_rows(sxy)
```

## Data statistics after track cut
```{r}
mdat %>% group_by(Animal) %>% summarize(n=n())
```

```{r}
###################################################
#filter by two whales for the moment
#mdat<-mdat %>% filter(Animal %in% c("131127","131136","131132","131133","154187","131134","131115","131111","131130","131116"))
mdat<-mdat %>% filter(Animal %in% c("131134"))

#remake map
troy<-makemap(mdat)
####################################################

#refactor animal
mdat$Animal<-as.numeric(as.factor(mdat$Animal))

##Split into format for JAGS

#total number of steps per track/animal
steps_all<-mdat %>% group_by(Animal,Track) %>% summarize(n=length(unique(step)))

# give each step a label
mdat<-mdat %>% group_by(Animal,Track,step) %>% mutate(jStep=1:n())

#Cast time array
j<-reshape2::acast(mdat,Animal~Track~step~jStep,value.var="j")

#how many observations per individual in each step
mdat$step<-factor(mdat$step,levels=1:max(steps_all$n))
idx<-reshape2::melt(table(mdat$Animal,mdat$Track,mdat$step))
colnames(idx)<-c("Animal","Track","step","jStep")
idx<-reshape2::acast(data=idx,Animal~Track~step)

#Individuals
ind=length(unique(mdat$Animal))

#tracks per indivudal
tracks<-mdat %>% group_by(Animal) %>% summarize(tracks=length(unique(Track))) %>% .$tracks

#steps per track
steps<-reshape2::acast(steps_all,Animal~Track,value.var="n")

#obs array
obs<-reshape2::melt(mdat,measure.vars=c("Longitude","Latitude"))
obs<-reshape2::acast(obs,Animal~Track~step~jStep~variable,fun.aggregate = mean)

#argos class array
mdat$argos.lc<-factor(mdat$Quality,levels=c(3,2,1,0,"A","B"))
mdat$numargos<-as.numeric(mdat$argos.lc)
obs_class<-reshape2::acast(mdat,Animal~Track~step~jStep,value.var="numargos",fun.aggregate = min)

obs_class[!is.finite(obs_class)]<-NA

#average dive depth array
maxdive<-reshape2::acast(mdat,Animal~Track~step~jStep,value.var="DepthMax",fun.aggregate = mean)

#fill the empty values
maxdive[!is.finite(maxdive)]<-NA

#average number of dives
diveduration<-reshape2::acast(mdat,Animal~Track~step~jStep,value.var="DurationMax")
```

## Dive Prior Shapes

### Dive Depth

```{r}
#Dives
travel_dive_prior<-abs(rnorm(1e5,rnorm(1e5,0,1/sqrt(0.0001)),1/sqrt(runif(1e5,0,500))))
forage_dive_prior<-abs(rnorm(1e5,rnorm(1e5,50,1/sqrt(0.0001)),1/sqrt(runif(1e5,0,500))))
rest_dive_prior<-abs(rnorm(1e5,rnorm(1e5,0,1/sqrt(0.01)),1/sqrt(runif(1e5,0,500))))
divepriors<-gather(data.frame(Travel=travel_dive_prior,Forage=forage_dive_prior,Resting=rest_dive_prior))
ggplot(divepriors) + geom_density(aes(x=value,fill=key),alpha=0.5) + theme_bw()
```

```{r,child="Bayesian/ThreeState.R",eval=T}
```

```{r,eval=T}
#source jags file
source("Bayesian/ThreeState.R")

#prior cov shape
R <- diag(c(1,1))
data=list(duration=diveduration,dive=maxdive,argos=obs,steps=steps,R=R,ind=ind,j=j,idx=idx,tracks=tracks,argos_class=obs_class)

#paramters to track
pt<-c("Traveling","Foraging","gamma","duration_mu","duration_tau","depth_mu","depth_tau","dive_new","state","E","Enew")

if(newModel){
  system.time(diving<-jags.parallel(model.file = "Bayesian/ThreeState.jags",data=data,n.chains=2,parameters.to.save=pt,n.iter=5000,n.burnin=4500,n.thin=2,DIC=FALSE))
  }

```

##Chains
```{r,eval=T}
#bind chains
pc_dive<-reshape2::melt(diving$BUGSoutput$sims.array)
colnames(pc_dive)<-c("Draw","chain","par","value")

#extract parameter name
pc_dive$parameter<-data.frame(str_match(pc_dive$par,"(\\w+)"))[,-1]

#Extract index
splitpc<-split(pc_dive,pc_dive$parameter)

#single index
splitpc[c("Traveling","Foraging","depth_mu","depth_tau","duration_mu","duration_tau")]<-lapply(splitpc[c("Traveling","Foraging","depth_mu","depth_tau","duration_mu","duration_tau")],function(x){
    sv<-data.frame(str_match(x$par,"(\\w+)\\[(\\d+)]"))[,3]
    pc<-data.frame(x,Behavior=sv)
    return(pc)
}) 


#3 index
splitpc[c("state")]<-lapply(splitpc[c("state")],function(x){
    sv<-data.frame(str_match(x$par,"(\\w+)\\[(\\d+),(\\d+),(\\d+)"))[,3:5]
    colnames(sv)<-c("Animal","Track","step")
    pc<-data.frame(x,sv)
    return(pc)
}) 

#4 index
splitpc[c("dive_new","E","Enew")]<-lapply(splitpc[c("dive_new","E","Enew")],function(x){
    sv<-data.frame(str_match(x$par,"(\\w+)\\[(\\d+),(\\d+),(\\d+),(\\d+)]"))[,3:6]
    colnames(sv)<-c("Animal","Track","step","jStep")
    pc<-data.frame(x,sv)
    return(pc)
}) 
#bind all matrices back together
pc_dive<-bind_rows(splitpc)
rm(splitpc)

ggplot(pc_dive[!pc_dive$parameter %in% c("dive_new","state","E","Enew","eval"),],aes(x=Draw,y=value,col=as.factor(chain))) + geom_line() + facet_wrap(~par,scales="free")
ggplot(pc_dive[!pc_dive$parameter %in% c("dive_new","state","E","Enew","eval"),],aes(x=value,fill=as.factor(chain))) + geom_histogram() + facet_wrap(~par,scales="free")

```

## Temporal autocorrelation in foraging 

Reconstructing the transition matrix

```{r,eval=F}
amatrix<-pc_dive %>% filter(parameter %in% c("Traveling","Foraging"))
amatrix<-amatrix %>%  dplyr::select(Draw,chain,parameter,Behavior,value) %>% spread(parameter,value) %>% mutate(Resting=1-(Traveling + Foraging))

tmatrix<-gather(amatrix,To,value,-Draw,-chain,-Behavior)

btable<-data.frame(Behavior=c(1,2,3),BState=c("Traveling","Foraging","Resting"))

tmatrix<-merge(tmatrix,btable)
head(tmatrix)
ggplot(tmatrix,aes(x=value)) + facet_grid(To~BState) + geom_histogram() 
ggplot(tmatrix,aes(x=value)) + facet_wrap(~To) + geom_density(aes(fill=BState)) 

#mean matrix?
#ggplot(tmatrix,aes(x=BState,y=To,fill=value)) + geom_tile() 
```

### Time foraging as a function of time

```{r,eval=F}
foraging_time<-function(alpha,draws,step){
  states<-c()
  
  #initial state is traveling
  states[1]<-(1)
  
  for(x in 2:draws){
    
    #probability of staying, or switching, to traveling state
    travel_prob=alpha %>% filter(Previous_State==states[x-1]) %>% sample_n(1) %>% .$value
    
    #pick next state
    states[x]<-rbinom(1,1,travel_prob)
  }
  
  #total hours of foraging
  total_foraging<-sum(states==0) * step
return(total_foraging)
}

states<-foraging_time(alpha=alpha,draw=20,step=12)

```

## Simulate dive depth
```{r}
depth_mu1<-pc_dive %>% filter(parameter %in% c("depth_mu"),Behavior==1)
depth_mu2<-pc_dive %>% filter(parameter %in% c("depth_mu"),Behavior==2)
depth_mu3<-pc_dive %>% filter(parameter %in% c("depth_mu"),Behavior==3)

depth_tau1<-pc_dive %>% filter(parameter %in% c("depth_tau"),Behavior==1)
depth_tau2<-pc_dive %>% filter(parameter %in% c("depth_tau"),Behavior==2)
depth_tau3<-pc_dive %>% filter(parameter %in% c("depth_tau"),Behavior==3)

#how many of each state to draw?
statecount<-pc_dive %>% filter(parameter=="state") %>% group_by(Animal,Track,step) %>% summarize(state=names(sort(table(value)))[1]) %>% ungroup() %>% mutate(Animal=as.numeric(as.character(Animal)),Track=as.numeric(as.character(Track))) %>% inner_join(mdat,by=c("Animal","Track","step")) %>% group_by(state) %>% summarize(n=n()) %>% data.frame()

#simulate dives 
travel_dives<-data.frame(MaxDepth=rtruncnorm(statecount[1,"n"], mean=depth_mu1$value,sd=sqrt(1/depth_tau1$value),a=0),Behavior="Traveling")
forage_dives<-data.frame(MaxDepth=rtruncnorm(statecount[2,"n"], mean=depth_mu2$value,sd=sqrt(1/depth_tau2$value),a=0),Behavior="Foraging")
rest_dives<-data.frame(MaxDepth=rtruncnorm(statecount[3,"n"], mean=depth_mu3$value,sd=sqrt(1/depth_tau3$value),a=0),Behavior="Resting")

pred_dives<-bind_rows(travel_dives,forage_dives,rest_dives)

ggplot(pred_dives)  + geom_histogram(data=mdat,aes(x=DepthMax),col="black")+ geom_histogram(alpha=0.8,aes(x=MaxDepth,fill=Behavior))  + theme_bw() + labs(x="Dive Depth (m)")
ggsave("Figures/DiveHist.jpg",height=4,width=7)

ggplot(pc_dive[pc_dive$parameter=="dive_new",])+  geom_density(aes(x=value),col='red') + geom_density(data=mdat,aes(x=DepthMax),col="black",size=1.5) + theme_bw() + labs(x="Dive Depth (m)")
ggsave("Figures/DivePredict.jpg",height=4,width=7)
```

## Simulate dive durations

```{r,eval=T}
duration_mu1<-pc_dive %>% filter(parameter %in% c("duration_mu"),Behavior==1)
duration_mu2<-pc_dive %>% filter(parameter %in% c("duration_mu"),Behavior==2)
duration_mu3<-pc_dive %>% filter(parameter %in% c("duration_mu"),Behavior==3)

duration_tau1<-pc_dive %>% filter(parameter %in% c("duration_tau"),Behavior==1)
duration_tau2<-pc_dive %>% filter(parameter %in% c("duration_tau"),Behavior==2)
duration_tau3<-pc_dive %>% filter(parameter %in% c("duration_tau"),Behavior==3)

#how many of each state to draw?
statecount<-pc_dive %>% filter(parameter=="state") %>% group_by(Animal,Track,step) %>% summarize(state=names(sort(table(value)))[1]) %>% ungroup() %>% mutate(Animal=as.numeric(as.character(Animal)),Track=as.numeric(as.character(Track))) %>% inner_join(mdat,by=c("Animal","Track","step")) %>% group_by(state) %>% summarize(n=n()) %>% data.frame()

#simulate dives 
travel_dives<-data.frame(DurationMax=rtruncnorm(statecount[1,"n"], mean=duration_mu1$value,sd=sqrt(1/duration_tau1$value),a=0),Behavior="Traveling")
forage_dives<-data.frame(DurationMax=rtruncnorm(statecount[2,"n"], mean=duration_mu2$value,sd=sqrt(1/duration_tau2$value),a=0),Behavior="Foraging")
rest_dives<-data.frame(DurationMax=rtruncnorm(statecount[2,"n"], mean=duration_mu3$value,sd=sqrt(1/duration_tau3$value),a=0),Behavior="Resting")

pred_dives<-bind_rows(travel_dives,forage_dives,rest_dives)

ggplot(pred_dives)  + geom_histogram(data=mdat,aes(x=DurationMax),col="black")+ geom_histogram(alpha=0.8,aes(x=DurationMax,fill=Behavior))  + theme_bw() 
ggsave("Figures/DurationHist.jpg",height=5,width=7)

ggplot(pred_dives)+  geom_density(aes(x=DurationMax),col='red') + geom_density(data=mdat,aes(x=DurationMax),col="black",size=1.5) + theme_bw() + labs(x="Dive Duration (min)")
ggsave("Figures/DurationPred.jpg",height=5,width=7)

```

# Temporal Variation in Dive Behavior

```{r}

#Take the most common estimate of behavior
Mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

#Combine posterior summary
mdat$Animal<-as.factor(mdat$Animal)
mdat$Track<-as.factor(mdat$Track)
mdat$step<-as.factor(mdat$step)

state_est<-pc_dive %>% group_by(Animal,Track,step) %>% filter(parameter=="state") %>% summarize(Behavior=Mode(value)) %>% inner_join(mdat)

state_est[state_est$Behavior==1,"Behavior"]<-"Traveling"
state_est[state_est$Behavior==2,"Behavior"]<-"Foraging"
state_est[state_est$Behavior==3,"Behavior"]<-"Resting"

```

```{r,fig.height=15,fig.width=12}
ggplot(state_est,aes(x=timestamp,y=-DepthMax,col=Behavior)) + geom_point() + geom_line(size=0.25,aes(group=1)) + facet_wrap(~Animal,scales="free",ncol=1) + theme_bw()
ggsave("Figures/TemporalBehavior.svg")
ggsave("Figures/TemporalBehavior.jpeg",height=8,width=12)

#by_state<-state_est %>% group_by(Animal,Track,step) %>% summarize(timestamp=min(timestamp),dive=mean(DepthMax),Behavior=unique(Behavior))
#ggplot(by_state,aes(x=timestamp,y=-dive,col=Behavior)) + geom_point() + geom_line(size=0.25,aes(group=1)) + facet_wrap(~Animal,scales="free",ncol=1) + theme_bw()
```

## Diel

```{r}
ggplot(data=state_est) + geom_boxplot(aes(x=LocalHour,y=DepthMax,fill=Behavior)) + ylab("Maximum Dive Depth (km)")+ labs(x="Hour (Local GMT+3)")
ggsave("Figures/Diel.svg")
ggsave("Figures/Diel.png",height=5,width=8,unit="in")

ggplot(data=state_est) + geom_boxplot(aes(x=LocalHour,y=DepthMax,fill=Behavior)) + ylab("Maximum Dive Depth (km)") + facet_wrap(~Month) + labs(x="Hour (Local GMT+3)")
ggsave("Figures/DielbyMonth.svg")
ggsave("Figures/DielbyMonth.png",height=5,width=10,unit="in")
```

## Month
```{r}
ggplot(data=state_est) + geom_boxplot(aes(x=Month,y=DepthMax,fill=Behavior)) + ylab("Maximum Dive Depth (km)")
ggsave("Figures/Month.svg")
ggsave("Figures/Month.png",height=5,width=8,unit="in")
```

```{r,fig.height=10,fig.width=10}
ggmap(troy) +geom_path(data=state_est,aes(x=Longitude, y=Latitude,group=paste(Animal,Track)),size=0.5) + geom_point(data=state_est,aes(x=Longitude, y=Latitude,col=Behavior),size=0.5) + mytheme
```

# Comparison with 2d movement model

* Movement model

```{r,eval=T}
#source jags file
source("Bayesian/Foraging.R")

#prior cov shape
R <- diag(c(1,1))
data=list(argos=obs,steps=steps,R=R,ind=ind,j=j,idx=idx,tracks=tracks,argos_class=obs_class)

#paramters to track
pt<-c("alpha","gamma","state")

if(newModel){
  system.time(foraging<-jags.parallel(model.file = "Bayesian/Foraging.jags",data=data,n.chains=2,parameters.to.save=pt,n.iter=10000,n.burnin=9500,n.thin=2,DIC=FALSE))
  }

```

##Chains

```{r,eval=T}
#bind chains
pc_dive2d<-reshape2::melt(foraging$BUGSoutput$sims.array)
colnames(pc_dive2d)<-c("Draw","chain","par","value")

#extract parameter name
pc_dive2d$parameter<-data.frame(str_match(pc_dive2d$par,"(\\w+)"))[,-1]

#Extract index
splitpc<-split(pc_dive2d,pc_dive2d$parameter)

#single index
splitpc[c("alpha")]<-lapply(splitpc[c("alpha")],function(x){
    sv<-data.frame(str_match(x$par,"(\\w+)\\[(\\d+)]"))[,3]
    pc<-data.frame(x,Behavior=sv)
    return(pc)
}) 

#3 index
splitpc[c("state")]<-lapply(splitpc[c("state")],function(x){
    sv<-data.frame(str_match(x$par,"(\\w+)\\[(\\d+),(\\d+),(\\d+)"))[,3:5]
    colnames(sv)<-c("Animal","Track","step")
    pc<-data.frame(x,sv)
    return(pc)
}) 

#bind all matrices back together
pc_dive2d<-bind_rows(splitpc)
rm(splitpc)

ggplot(pc_dive2d[!pc_dive2d$parameter %in% c("state"),],aes(x=Draw,y=value,col=as.factor(chain))) + geom_line() + facet_wrap(~par,scales="free")
```

```{r}

#Take the most common estimate of behavior
Mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

#Combine posterior summary
mdat$Animal<-as.factor(mdat$Animal)
mdat$Track<-as.factor(mdat$Track)
mdat$step<-as.factor(mdat$step)

state_est2d<-pc_dive2d %>% group_by(Animal,Track,step) %>% filter(parameter=="state") %>% summarize(Behavior =Mode(value)) %>% inner_join(mdat)

state_est2d[state_est2d$Behavior==1,"Behavior"]<-"Traveling"
state_est2d[state_est2d$Behavior==2,"Behavior"]<-"Foraging"

state_est$Model<-"3D"
state_est2d$Model<-"2D"

allmodels<-bind_rows(state_est,state_est2d)

```

```{r}
ggmap(troy) +geom_path(data=allmodels,aes(x=Longitude, y=Latitude,group=paste(Animal,Track)),size=0.5) + geom_point(data=allmodels,aes(x=Longitude, y=Latitude,col=Behavior),size=0.5) + theme_bw() + mytheme  + facet_wrap(~Model)
ggsave("Figures/SpatialPred.jpg",height=4,width=7)
```

Where does the 3d predict foraging that the 2d misses?

```{r}
allmodels<-spread(allmodels,Model,Behavior)
allmodels$Mismatch<-F
allmodels[allmodels$`2D`=="Traveling" & allmodels$`3D` %in% c("Foraging"),"Mismatch"]<-TRUE

ggmap(troy)+geom_path(data=allmodels,aes(x=Longitude, y=Latitude,group=paste(Animal,Track)),size=.25) + geom_point(data=allmodels,aes(x=Longitude, y=Latitude,col=Mismatch)) + theme_bw() + mytheme + scale_color_manual(values=c("Black","Red"))
ggsave("Figures/Mismatch1.jpg",height=4,width=7)
```

Where does the 3d predict resting that the 2d misses?

```{r}
allmodels$Mismatch<-F
allmodels[allmodels$`2D`=="Traveling" & allmodels$`3D` %in% c("Resting"),"Mismatch"]<-TRUE

ggmap(troy)+geom_path(data=allmodels,aes(x=Longitude, y=Latitude,group=paste(Animal,Track)),size=.25) + geom_point(data=allmodels,aes(x=Longitude, y=Latitude,col=Mismatch)) + theme_bw() + mytheme + scale_color_manual(values=c("Black","Red"))
ggsave("Figures/Mismatch1.jpg",height=4,width=7)
```

Where does the 2d predict foraging that the 3d refines to traveling?

```{r}
allmodels$Mismatch<-F
allmodels[allmodels$`2D`=="Foraging" & allmodels$`3D`=="Traveling","Mismatch"]<-TRUE

ggmap(troy) +geom_path(data=allmodels,aes(x=Longitude, y=Latitude,group=paste(Animal,Track)),size=.25) + geom_point(data=allmodels,aes(x=Longitude, y=Latitude,col=Mismatch)) + mytheme + scale_color_manual(values=c("Black","Red"))
ggsave("Figures/Mismatch2.jpg",height=4,width=7)
```

Where does the 2d predict foraging that the 3d refines to resting?

```{r}
allmodels$Mismatch<-F
allmodels[allmodels$`2D`=="Foraging" & allmodels$`3D`=="Resting","Mismatch"]<-TRUE

ggmap(troy)+geom_path(data=allmodels,aes(x=Longitude, y=Latitude,group=paste(Animal,Track)),size=.25) + geom_point(data=allmodels,aes(x=Longitude, y=Latitude,col=Mismatch)) + theme_bw() + mytheme + scale_color_manual(values=c("Black","Red"))
ggsave("Figures/Mismatch2.jpg",height=4,width=7)
```

## Posterior Checks

The goodness of fit is a measured as chi-squared. The expected value is compared to the observed value of the actual data. In addition, a replicate dataset is generated from the posterior predicted intensity. Better fitting models will have lower discrepancy values and be 
Better fitting models are smaller values and closer to the 1:1 line. A perfect model would be 0 discrepancy. This is unrealsitic given the stochasticity in the sampling processes. Rather, its better to focus on relative discrepancy. In addition, a model with 0 discrepancy would likely be seriously overfit and have little to no predictive power.

```{r,fig.height=4,fig.width=8,message=F,warning=F}

fitstat<-pc_dive %>% filter(parameter %in% c("E","Enew")) %>% group_by(parameter,Draw,chain) %>% summarize(fit=mean(value))

fitstat<-dcast(fitstat,Draw+chain~parameter,value.var="fit")

ymin<-min(c(fitstat$E,fitstat$Enew)) - min(c(fitstat$E,fitstat$Enew)) * .1
ymax<-max(c(fitstat$E,fitstat$Enew)) + max(c(fitstat$E,fitstat$Enew)) * .1
ggplot(fitstat,aes(x=E,y=Enew)) + geom_point() + theme_bw() + labs(x="Discrepancy of observed data",y="Discrepancy of replicated data") + geom_abline() + coord_fixed() + ylim(ymin=ymin,ymax=ymax) + xlim(xmin=ymin,xmax=ymax)
ggsave("Figures/PosteriorCheck.jpg",height = 4,width = 4)
fitstat %>% group_by() %>% summarize(mean(E),var(Enew))
```

```{r}
save.image("/Users/Ben/Dropbox/Whales/Dive/WhalePhys.RData")
```