---
title: "Foraging Simulation"
author: "Ben Weinstein"
date: "`r Sys.time()`"
output: 
  html_document:
    toc: true
    number_sections: true
    theme: united
---

#Parameters

## Empirical parameters estimated from whale tagging data

Behavior allocation (Foraging,Traveling) at each time step 

$$state_t \sim(Binomial(\alpha)) $$

Average maximum depth 

$$Normal(\mu = dive_\mu, sd= \frac{1}{dive_\tau}) $$

Average number of dives 

$$Poisson(\lambda = dive_{count})$$

## Parameters estimated from literature 

Initial body mass (Tyson et al. 2016, Lockyer 1976)

$$M = 39614 kg $$

Metabolic rate while traveling Tyson et al. (2016),Williams et al. (2001).

$$m_{traveling} = -2.32 \frac{KJ}{sec} $$

Metabolic rate while foraging Tyson et al. (2016),Williams et al. (2001).

$$m_{foraging} = -4.14 \frac{KJ}{sec}$$

Lunges at depth, Ware 2011 reported ~40 sec interlunge time, with 1.5m/s dive speed. 

$$Lunges \sim Poisson(\lambda=\frac{\text{dive_depth}}{50})$$

Energetic cost per lunge (Tyson et al. 2016)

$$E_{lunge} = -298.08 KJ$$

**this seems widly too low** 

Energetic gain per lunge 41KJ/sec (Tyson et al 2016) and 15.5 sec per lunge (Ware et al. 2011)

$$41 \frac{KJ}{sec} * 15.5 \frac{sec}{lunge} = 635 \frac{KJ}{Lunge}$$

Metabolic efficiency, energy lost to extretion (Bejarano et al. 2017)

$$ Beta(43.60,5.71) $$

```{r,echo=F,warning=F,message=F}
library(dplyr)
library(ggplot2)
library(truncnorm)
library(knitr)
opts_chunk$set(echo=F,warning=F,message=F,fig.width=11)
```

```{r}
#encode foraging parameters
pars<-list(
  iterations=20,## run simulation loop
  draws=120, ## number of 6 hour periods
  t_interval=6, ## time interval in hours,
  alpha=read.csv("Data/Alpha.csv"), ##empirical transition probability
  depth_mu2 = 0.22 , ## Average dive depth
  depth_tau2 = sqrt(1/92) ,## variance in dive depth
  lambda_count2=4.5, ## average number of dives during foraging
  lgain=635, ## KJ gain per lunge
  lcost=-298, ## KJ cost per lunche
  lungef=50*1000, ## Number of lunges per X km depth
  mass= 39614, ## initial mass of whale
  metabolic_efficiency=rbeta(1,43.60,5.71), ## % efficiency of metabolic action (loss to extretion)
  metabolic_foraging = -4.14, ## KJ cost of foraging per second
  metabolic_traveling=-2.32 ##KJ/second
  
  )
```

```{r}
##########
#Functions
##########

behavior<-function(pars){
  states<-c()
  
  #initial state is traveling
  states[1]<-(1)
  
  for(x in 2:pars$draws){
    
    #probability of staying, or switching, to traveling state
    travel_prob=pars$alpha %>% filter(Previous_State==states[x-1]) %>% sample_n(1) %>% .$value
    
    #pick next state
    states[x]<-rbinom(1,1,travel_prob)
  }
  
  #total hours of foraging
return(states)
}
```

```{r}
#foraging energy functions
sim_lunges<-function(divedepth){
  rpois(1,divedepth * pars$lungef)
}

sim_lcost<-function(lunges){
  return(lunges*pars$lcost)
}

sim_lgain<-function(lunges){
  return(lunges *pars$lgain )
}
```


```{r}
#Traveling state function

traveling_state<-function(pars){
    
    #Cost of traveling
    energy<-pars$metabolic_traveling * pars$t_interval * 3600 ##metabolic cost of traveling 
    
  return(energy)
  }
 
```

```{r}

#Foraging state function

foraging_state<-function(pars){
    divecount<-rpois(1,pars$lambda_count2)
    
    dive_energy<-c()
    for(dive in 1:divecount){
      
      #dive length
      divedepth<-rtruncnorm(1,pars$depth_mu2,pars$depth_tau2,a=0.01)  
      
      #lunges 
      lunges<-sim_lunges(divedepth)
      
      #gain per lunge
      gain_lunge<-sim_lgain(lunges)
      
      #costs per lunche
      cost_lunge<-sim_lcost(lunges)
      
      dive_energy[x]<-gain_lunge-cost_lunge
    }
    
    #sum across dives, subtract foraging metabolic rate
    energy<-sum(dive_energy,na.rm=T) - (pars$metabolic_foraging * pars$t_interval*3600)
     return(energy)
  }
 
```

```{r}
simulation<-function(pars){
  
  simid<-runif(1,0,10000000)
  
  #foraging states in a given time interval
  states<-behavior(pars=pars)
  
  energy<-c()
  #time foraging periods
  for (x in 1:length(states)){
    
    #if foraging
    if(states[x]==1){
      energy[x]<-foraging_state(pars)
    } else{
      energy[x]<-traveling_state(pars)
    }
  }
  
  #Convert energy to calories
  energy<-energy * pars$metabolic_efficiency
  
  #cumulative energy
  result<-data.frame(ID=simid,step=1:length(states),state=states,energy=energy)
  result<-result %>% mutate(cumulative_energy=cumsum(energy))
  
  #time
  result$days<-result$step * pars$t_interval/24
  
  return(result)
}
```

```{r}

results<-list()
for(x in 1:pars$iterations){
  results[[x]]<-simulation(pars)
}
results<-bind_rows(results)
```

Raw results
```{r}
ggplot(results,aes(x=step,y=cumulative_energy,group=as.factor(ID))) + geom_point() + geom_line() + theme_bw() + labs(y="Cumulative KJ Gain")
```

Scaled by time
```{r}
ggplot(results,aes(x=days,y=cumulative_energy,group=as.factor(ID))) + geom_point() + geom_line() + theme_bw()+ labs(y="Cumulative KJ Gain")
```

# Questions and Improvements

* How to convert from energy to grams? Weight of blubber?
* Need to improve the dive count mixture distribution
* Include dive duration, where would it help inference?
* Are we double containing metabolic costs of foraging and costs of lunges?
* Mass specific effects?
* Parameter bounds for all variables

# Future simulation directions

* Variable krill swarm size

* Variable krill quality

* Temporal variation in foraging rates, depths

* Explicitely spatial structure, movement, krill swarms, etc.
 
* Sensitivity